---
title: "Practical Machine Learning Course Project"
author: "Adithya Diddapur"
date: "2 April 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Machine Learning Course Project

###Part 1: Exploratory Data Analysis
Let's start by loading the data and examining it for missing values. We are going to examine it for missing values by producing a missmap plot from the Amelia package. This plot gives us a good understanding of
which variables are missing from the data.
```{r, cache=TRUE, message=FALSE}
trainingData <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
testingData <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

library(Amelia)
missmap(trainingData, main="Missing Values")
```
As you can see, around halof of the variables have most observations, but some of the variables are
missing most observations. If we include them, they'll increase the bias in our model, so lets remove
them by filtering out variables where more than 50% of the observations are missing.
```{r, cache=TRUE}
variablesToRemove <- sapply(trainingData, function(x)sum(is.na(x))>(length(x)*0.5))
filteredTrainingData <- trainingData[!variablesToRemove]
```

###Part 2: Model Selection
Let's start by creating a linear model, random forest, and lda model and seeing how well they performs.
```{r, cache=TRUE}
library(caret)
lModel <- lm(classe~., filteredTrainingData)
confusionMatrix()
```

